---
title: "LAGOS Spatial Analysis"
author: "AJ Lee"
date: "2/21/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(ggthemes)
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)

```{r data-read}
# #Lagos download script
LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())


#Load in lagos
lagos <- lagosne_load()

#Grab the lake centroid info
lake_centers <- lagos$locus


```



### Convert to spatial data
```{r}
#Look at the column names
#names(lake_centers)

#Look at the structure
#str(lake_centers)

#View the full dataset
#View(lake_centers %>% slice(1:100))

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100) 

subset_baser <- spatial_lakes[1:100,]

#Dynamic mapviewer
#mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)

minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```



# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r}

Iowa_Illinois <- states %>% 
              filter(name %in% c('Iowa', 'Illinois')) %>% 
              st_transform(2163)

mapview(Iowa_Illinois)

```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa
combined? How does this compare to Minnesota?

```{r}
Iowa_Illinois_lakes <- spatial_lakes[Iowa_Illinois,]
nrow(Iowa_Illinois_lakes)
nrow(minnesota_lakes)

```


## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

```{r}
# Make Iowa's lakes size data 
Iowa <- states %>% 
       filter(name=='Iowa') %>% 
       st_transform(2163)

Iowa_lakes <- spatial_lakes[Iowa,]

Iowa_lakes_size <- data.frame(Iowa_lakes$lake_area_ha) %>% 
                   rename(Lake_Size=1) %>% 
                   mutate(State='Iowa')

  
# Make Minnesota's lakes size data
minnesota_lakes_size <- data.frame(minnesota_lakes$lake_area_ha) %>%
                        rename(Lake_Size=1) %>% 
                        mutate(State='Minnesota')


# Bind two data frames
Compare_lake_size <- rbind(Iowa_lakes_size, minnesota_lakes_size)

ggplot(Compare_lake_size, aes(x=Lake_Size, fill=State)) +
       xlim(1,200) +
       xlab('Lake Size(ha)') +
       geom_histogram(bins = 300, alpha=0.6, position = "identity") +
       theme_few() + 
       scale_color_few() + 
       theme(legend.position=c(0.8,0.8))

```


## 4) Make an interactive plot of lakes in Iowa and Illinois and color them
by lake area in hectares

```{r}
Iowa_Illinois_lakes %>%
  arrange(-lake_area_ha) %>%
  slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')


```


## 5) What other data sources might we use to understand how reservoirs and 
natural lakes vary in size in these three states? 
```{r}
Compare_lake_length <- rbind(minnesota_lakes, Iowa_Illinois_lakes)

Compare_lake_length %>% 
  arrange(-lake_perim_meters) %>% 
  slice(1:1000) %>% 
  mapview(., zcol='lake_perim_meters')

```



