---
title: "Lake Water Quality Analysis"
author: "AJ LEE"
date: "2/21/2022"
output: html_document
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(lubridate) #For dealing with date and time
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)
```{r data-read, warning=F}
#Lagos download script
# lagosne_get(dest_folder = LAGOSNE:::lagos_path(),overwrite=T)

#Load in lagos
lagos <- lagosne_load()


#Grab the lake centroid info
lake_centers <- lagos$locus

# Make an sf object 
spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>% 
                          st_transform(2163)

#Grab the water quality data
nutr <- lagos$epi_nutr

#Look at column names
#names(nutr)

```

### Subset columns nutr to only keep key info that we want


```{r}
clarity_only <- nutr %>%
  select(lagoslakeid,sampledate,chla,doc,secchi) %>%
  mutate(sampledate = as.character(sampledate) %>% ymd(.))

```


### Keep sites with at least 200 observations 

```{r}

#Look at the number of rows of dataset
#nrow(clarity_only)

chla_secchi <- clarity_only %>%
  filter(!is.na(chla),
         !is.na(secchi))

# How many observatiosn did we lose?
# nrow(clarity_only) - nrow(chla_secchi)


# Keep only the lakes with at least 200 observations of secchi and chla
chla_secchi_200 <- chla_secchi %>%
  group_by(lagoslakeid) %>%
  mutate(count = n()) %>%
  filter(count > 200)

```


### Join water quality data to spatial data

```{r}
spatial_200 <- inner_join(spatial_lakes,chla_secchi_200 %>%
                          distinct(lagoslakeid,.keep_all=T),
                          by='lagoslakeid')

# mapview(spatial_200)

```

### Mean Chl_a map

```{r}
### Take the mean chl_a and secchi by lake

mean_values_200 <- chla_secchi_200 %>%
  # Take summary by lake id
  group_by(lagoslakeid) %>%
  # take mean chl_a per lake id
  summarize(mean_chl = mean(chla,na.rm=T),
            mean_secchi=mean(secchi,na.rm=T)) %>%
  #Get rid of NAs
  filter(!is.na(mean_chl),
         !is.na(mean_secchi)) %>%
  # Take the log base 10 of the mean_chl
  mutate(log10_mean_chl = log10(mean_chl))

#Join datasets
mean_spatial <- inner_join(spatial_lakes,mean_values_200,
                          by='lagoslakeid')  
                 
#Make a map
mapview(mean_spatial,zcol='log10_mean_chl')


```


# Class work

## 1) What is the correlation between Secchi Disk Depth and Chlorophyll a for sites with at least 200 observations?

- Here, I just want a plot of chl vs secchi for all sites  

<span style="color:blue"> 
**A: The plot below shows that high chlorophyll(chla) shorten the depth of secchi disk, which means a negative correlation.** </span>

```{r}
#Your code here
plot(mean_chl ~ mean_secchi, data=mean_spatial)

```


## Why might this be the case? 
<span style="color:blue">
**A: The increase in chlorophyll in the water means high turbidity, and it is difficult to see the secchi disk to the depths because it interferes with visibility in the water. ** </span>


## 2) What states have the most data? 

### 2a) First you will need to make a lagos spatial dataset that has the total number of counts per site.

```{r}
#head(chla_secchi)

chla_secchi_most <- chla_secchi %>%
  group_by(lagoslakeid) %>%
  mutate(count = n()) %>% 
  summarize(Total_by_lake = sum(count))

spatial_most <- inner_join(spatial_lakes,chla_secchi_most,
                          by='lagoslakeid')

#mapview(spatial_most)

```


### 2b) Second, you will need to join this point dataset to the us_boundaries data. 

```{r}
states <- us_states() %>% 
          st_transform(2163)

US_spatial_most <- st_join(spatial_most, states)

#mapview(US_spatial_most)

```


### 2c) Then you will want to group by state and sum all the observations in that state and arrange that data from most to least total observations per state. 

<span style="color:blue"> 
**A: The state has the most data about lake is Minnesota, which observation number is about over 6 million in chlorophyll and secchi data record.** </span>

```{r}
US_lake_Observation <- US_spatial_most %>% 
                       group_by(name) %>% 
                       summarize(Observation=sum(Total_by_lake)) %>% 
                       arrange(-Observation)
head(US_lake_Observation)

```

## 3) Is there a spatial pattern in Secchi disk depth for lakes with at least 200 observations?  

<span style="color:blue">
**A: It can be seen that the high secchi depth value, which means the clarity of water, is mainly distributed in the northeastern regions such as Minnesota, Michigan, New York, and Maine.**</span> 
```{r}

# Extract the secchi data from original nutrition data
Secchi <- nutr %>% 
          select(lagoslakeid, sampledate, secchi) %>% 
          mutate(sampledate = as.character(sampledate) %>% ymd(.))

# Filter only the data above 200 observation
Secchi_200 <- Secchi %>% 
              group_by(lagoslakeid) %>% 
              mutate(count=n()) %>% 
              filter(count > 200)

# Join the data to spatial platform
Secchi_200_spatial <- inner_join(spatial_lakes, Secchi_200, by='lagoslakeid') %>% 
                      arrange(-secchi) %>% 
                      slice(1:30000)

# Check a spatial pattern using mapview 
mapview(Secchi_200_spatial, zcol='secchi')

```


